Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-01 paper/ms.tex
updating the methodology

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Finding the mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$: the importance of cosmic variance}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a unique and narrow set of values
for the parameters in the model. In particular, it is always possible
to find a model that is consistent with observations and has any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq 1.0$
.  Nevertheless, in the models we can define three different families
of parameters based on their mass range, $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$, and occupation fraction,
$f_{\rm occ}$. The dominant family is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm
  occ}\leq 0.3$ and a maximum mass $M_{\rm  max}< 10^{12}$\hMsun.  The
existence of this dominant family gives support to the idea that the
most massive dark matter halos at that epoch do not host the brightest
LAEs and that only a small fraction of star forming galaxies can be
actually detected as LAEs.

\end{abstract}

@@

Our method is based on the comparison of observatios and mock
catalogs on the surface number density and the angular correlation
function of LAEs.  In the next subsections we describe in detail the four key
elements of this workflow. First, we present the observations we take
as a benchmark. Second, the N-body simulation and the halo catalogs we
use. Third, the simplified model that allows us to translated halo
catalogs into mock LAE observations. Fourth, the statistics we use to
compare observational results against our theoretical predictions.

@@

One of the main observational benchmarks we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 subfields, 
this corresponds to average surface number density of $0.20\pm 0.01$ arcmin$^{-2}$.

@@

The second constraint is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement. Instead
we use the results by \cite{Hayashino2004} and
\cite{Ouchi2008,Ouchi2010}.

\cite{Hayashino2004} observed in the densest field of
SSA22 while \cite{Ouchi2008} observed on a 1 deg$^2$ sky of the
Subaru/XMM-Newton Deep Survey (SXDS) Field.  There are some
differences between these observations and those iby
\cite{Yamada2012}. The details in the color selection, corresponding
limiting luminosities and EW thresholds are different in these
references. Nevertheless we use these observations as additional
constraints in spite of the fact that the first selected models are
based only on the surface density statistics by \citep{Yamada2012}.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. We focus our work on halos more massive
than $1\times 10^{10}$\hMsun that are resolved with at least $70$
particles, the reasons for this choice are explained in the next
section.

@@

In our model a dark matter halo can only host one LAE at most. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that host a \emph{detectable LAE}. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two things: the presence of a star
forming galaxy in a halo and its detectability as a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ and $f_{\rm occ}$.

For each model ${\mathcal M}$ we create a set of mock fields from
disjoint volumes in the simulation. Each volume has the same geometry
probed by Suprime-CAM and the narrow band filter, namely rectangular
cuboids of dimensions $46\times 35\times 41$ $h^{-3}$Mpc$^{3}$ where
the last dimension goes in the redshift direction. This corresponds to
a total area of $880$ arcmin$^{2}$ in each mock field. We construct a total
$5\times 7 \times 6=210$ of such volumes from a snapshot in the Bolshoi
simulation. In each mock field a LAE is assigned to the position of a
dark matter halo if the halo mass is in the range allowed by the model
$M_{\rm min}<M_{\rm h}<M_{\rm max}$ and a random variable taken from
an homogeneous distribution $0\leq \xi<1$ is smaller than the occupation
fraction $\xi<f_{\rm occ}$.

Next we construct mock surveys by making groups of $11$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. The grouping is done in two
different ways. In the first way, called {\texttt match}, we follow
the clustering of the observed fields. From the $11$ mock fields, $7$
are constructed from contiguous fields in the simulation to mimic the
SSA22 region, $3$ are also contiguous between them but not to the first
$7$ fields to mimic the SXDS fields and finally $2$ non-contiguous
fields to imitate the SDF and GOODS-North field. Our main goal with
this selection is to test the impact on the final statistics of having
$7$ clustered fields. The second way to group the mock fields is called {\texttt
  random}, whereby all the $11$ fields are selected in such a way as
to avoid that any two volumes are contiguous. In this \documentname we
only report the results obtained by the {\texttt match} method that
are also obtained in the {\texttt random} selection.

\subsection{Sampling and Selecting Models}

@@

The lower limit for the parameter $M_{\rm min}$ is set by the minium
occupation fraction we are able to consider. At $M_{\rm
  min}=10^{10}$\hMsun the halo number density is already $\sim 10$
times higher than the observational constraints for LAEs. This means
that models in that mass range and an occupation fraction $f_{\rm
  occ}$ have the possibility to be compatible with observations. Lower
values for $M_{\rm min}$ require $f_{\rm occ}<0.1$, which are not
considered in this \documentname.

For each mock survey generated in a given model ${\mathcal M}$ we
compute the surface density in the $12$ mock fields. We perform a
Kolmogorov-Smirnov (KS) to compare this mock date against the $12$
observational values. From this test we obtain a value $0<P<1$ to
reject the null hypothesis, namely that two data sets come from the
same distribution. In this paper we consider that for values $P>0.05$
the two distributions can be thought as coming from the same
distribution.

@@

We then calculate the  angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Davis \& Peebles  estimator \citep{Davis1983} ) only on the densest
subfield in all the 15 mock surveys corresponding to the SSA22
region. We then take the mean ACF over the 15  mock surveys  for
each model as well as its the standard deviation to include
the effect of cosmic variance.

>>>>

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a unique and narrow set of values
for the parameters in the model. In particular, it is always possible
to find a model that is consistent with observations and has any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq 1.0$
.  Nevertheless, in the models we can define three different families
of parameters based on their mass range, $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$, and occupation fraction,
$f_{\rm occ}$. The dominant family is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm
  occ}\leq 0.3$ and a maximum mass $M_{\rm  max}< 10^{12}$\hMsun.  The
existence of this dominant family gives support to the idea that the
most massive dark matter halos at that epoch do not host the brightest
LAEs and that only a small fraction of star forming galaxies can be
actually detected as LAEs.
\end{abstract}

@@

Our method is based on the comparison of observations and mock
catalogs. We use two different kinds of statistics to perform the
comparison: (i) the distribution of the surface number density
across fields fields and (ii) the angular correlation function measured in
some fields.

In the next subsections we describe in detail the four key
elements of this workflow. First, we present the observations we take
as a benchmark. Second, we describe the main characteristics of the
N-body simulation and the halo catalogs we use. Third, we recount the
important parameters of the simplified model that we use to populate
the halo catalogs with LAEs. Finally we describe some of the
statistical tests we adopt to compare observations and mocks.

@@

The first observational constraint we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter used in the survey is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 subfields, 
this corresponds to average surface number density of $0.20\pm 0.01$
arcmin$^{-2}$.

@@

The second constraint is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement. Instead
we use the results by \cite{Hayashino2004} and
\cite{Ouchi2008,Ouchi2010}. \cite{Hayashino2004} observed in the
densest field of SSA22 while \cite{Ouchi2008} observed on a 1 deg$^2$
sky of the SXDS Field.  There are some differences between these
observations and those by \cite{Yamada2012}. The details in the color
selection, corresponding limiting luminosities and EW thresholds are
different in these references. Nevertheless we use these observations
as additional constraints in spite of the fact that the first selected
models are based only on the surface density statistics by
\citep{Yamada2012}.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. For each halo in the box we store its
position in the box (3-D coordinates) and FOF mass. We focus our work
on halos more massive than $1\times 10^{10}$\hMsun that are resolved
with at least $70$ particles, the reasons for this choice are
explained in the next sub-section.

@@

We assume that a dark matter halo can only host one
detectable LAE at most.  There are three parameters that
decide whether a halo host a LAE: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that host a detectable LAE. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two phenomena: the presence of a star
forming galaxy in a halo and its detectability as a LAE.  We do not
peform a explicit modeling for LAE detectability. Our model does not
assign a luminosity or escape fraction to each LAE.  We are only
interested in constraining the halo mass range hosting detectable LAEs
under the conditions defined by \cite{Yamada2012}.

In what follows we note by the letter ${\mathcal M}$ a model
defined by a particular choice of the three scalar parameters $M_{\rm
  min}$, $M_{\rm  max}$ and $f_{\rm occ}$. For each model ${\mathcal
  M}$ we create a set of mock fields from disjoint volumes in the
simulation. Each volume has the same geometry probed by Suprime-CAM
and the narrow band filter, namely rectangular cuboids of dimensions
$46\times 35\times 41$ $h^{-3}$ Mpc$^{3}$ where the last dimension goes
in the redshift direction. This corresponds to a total area of $880$
arcmin$^{2}$ in each mock field. We construct a total $5\times 7
\times 6=210$ of such volumes from a snapshot in the Bolshoi
simulation. In each mock field a LAE is assigned to the position of a
dark matter halo if the halo mass is in the range allowed by the model
$M_{\rm min}<M_{\rm h}<M_{\rm max}$ and a random variable taken from
an homogeneous distribution $0\leq \xi<1$ is smaller than the occupation
fraction $\xi<f_{\rm occ}$.

Next we construct mock surveys by making groups of $12$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. The grouping of the $11$
mock fields into a mock catalog is done in two different ways. The
first is called {\texttt match}, it follows the clustering of the
observational fields. From the $12$ mock fields, $7$ are constructed
from contiguous fields in the simulation to mimic the SSA22 region,
$3$ are also contiguous between them but not to the first $7$ fields
to mimic the SXDS fields and finally $2$ non-contiguous fields to
imitate the SDF and GOODS-North field.  The second way to group the
mock fields is called {\texttt   random}, whereby all the $11$ fields
are selected in such a way as to avoid that any two volumes are
contiguous. In this \documentname we only report the results obtained
by the {\texttt match} method and mention explicitly differences
observed with the {\texttt random} selection.

\subsection{Exploring and Selecting Good Models}

@@

The lower limit for the parameter $M_{\rm min}$ is set by the minium
occupation fraction we are able to consider. At $M_{\rm
  min}=10^{10}$\hMsun the halo number density is already $\sim 10$
times higher than the observational constraints for LAEs. This means
that models in that mass range and an occupation fraction $f_{\rm
  occ}=0.1$ have the possibility to be compatible with observations. Lower
values for $M_{\rm min}$ require $f_{\rm occ}<0.1$, which are not
considered in this \documentname.

For each mock survey generated in a given model ${\mathcal M}$ we
compute the surface density in the $12$ mock fields. We perform a
Kolmogorov-Smirnov (KS) to compare this mock data against the $12$
observational values. From this test we obtain a value $0<P<1$ to
reject the null hypothesis, namely that two data sets come from the
same distribution. In this paper we consider that for values $P>0.05$
the two distributions can be thought as coming from the same
distribution.

@@

The second constraint  comes from the angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Davis \& Peebles  estimator \citep{Davis1983} ) only on the densest
subfield in all the 15 mock surveys corresponding to the SSA22
region. We then take the mean ACF over the 15  mock surveys  for
each model as well as its the standard deviation to include
the effect of cosmic variance.
